\begin{center}
%\section*{\MakeUppercase{ОСНОВНИЙ ЗМІСТ РОБОТИ}}
\textbf{\MakeUppercase{\fullbibtitle}}
\end{center}

\begin{enumerate}[label=\arabic*$^*$.,leftmargin=0em,itemindent=3em]
\item
Explanation of commonly observed shunt currents in c-Si solar cells by means of
  recombination statistics beyond the Shockley-Read-Hall approximation~/
  Silke~Steingrube, Otwin~Breitenstein, Klaus~Ramspeck et~al.~// \emph{J.
  Appl. Phys.} ---
  2011. --- July. ---
  Vol. 110, no.~1. ---
  P.~014515.



\item
\emph{Gopal,~Vishnu}. Contribution of Dislocations to the Zero-Bias
  Resistance-Area Product of {LWIR} {H}g{C}d{T}e Photodiodes at Low
  Temperatures~/ Vishnu~Gopal, Sudha~Gupta~// \emph{IEEE Trans. Electron
  Devices}. ---
  2004. --- Jul. ---
  Vol.~51, no.~7. ---
  P.~1078--1083.
  
\item
\emph{Mirzade,~Fikret}. Elastic wave propagation in a solid layer with
  laser-induced point defects~/ Fikret~Mirzade~// \emph{J. Appl. Phys.} ---
  2011. --- Sep. ---
  Vol. 110, no.~6. ---
  P.~064906.  

\item
Определение параметров глубоких уровней
  по дифференциальным коэффициентам
  вольт--амперных характеристик~/
  С.В.~Булярский, М.О.~Воробьев, Н.С.~Грушко,
  А.В.~Лакалин~// \emph{Письма в журнал
  технической физики}. ---
  1999. ---
  Т.~25, {№}~5. ---
  {С.}~22--27.

\item
\emph{Tung,~Raymond~T.} Recent advances in {S}chottky barrier concept~/
  Raymond~T.~Tung~// \emph{Materials Science and Engineering: R: Reports}.
  ---
  2001. --- Nov. ---
  Vol.~35, no. 1--3. ---
  P.~1--138.

\item
Distinction between the {P}oole-{F}renkel and tunneling models of
  electric-field-stimulated carrier emission from deep levels in
  semiconductors~/ S.~D.~Ganichev, E.~Ziemann, W.~Prettl et~al.~//
  \emph{Phys. Rev. B}. ---
  2000. --- Apr. ---
  Vol.~61, no.~15. ---
  P.~10361--10365.

\item
Schottky Barrier Height Inhomogeneity of Ti/n-GaAs Contact Studied by the I-V-T
  Technique~/ Yu-Long~Jiang, Guo-Ping~Ru, Fang~Lu et~al.~// \emph{Chin.
  Phys. Lett.} ---
  2002. --- Apr. ---
  Vol.~19, no.~4. ---
  P.~553--556.

\item
\emph{Brailsford,~A.~D.} Abrupt--Kink Model of Dislocation Motion~/
  A.~D.~Brailsford~// \emph{Phys. Rev.} ---
  1961. --- May. ---
  Vol. 122, no.~3. ---
  P.~778--786.

\item
\emph{Pipinys,~P.} Temperature dependence of reverse--bias leakage current
  in {G}a{N} {S}chottky diodes as a consequence of phonon--assisted tunneling~/
  P.~Pipinys, V.~Lapeika~// \emph{J. Appl. Phys.} ---
  2006. --- May. ---
  Vol.~99, no.~9. ---
  P.~093709.

\item
\emph{Jafar,~M M Abdul-Gader}. High-bias current--voltage--temperature
  characteristics of undoped rf magnetron sputter deposited boron carbide
  ({B}$_5${C})/p--type crystalline silicon heterojunctions~/ M~M
  Abdul-Gader~Jafar~// \emph{Semicond. Sci. Technol.} ---
  2003. --- Jan. ---
  Vol.~18, no.~1. ---
  P.~7--22.
\end{enumerate}
